Robot Team III: Scotty Strikes Back

### 2024-04-22 paper/paper.tex

Inland water bodies pose a unique challenge for analysis by remote sensing imagery due to their complex spectral characteristics and small-scale spatial variability; nevertheless, significant effort has gone into the development of techniques for estimating optically-active water quality parameters such as chlorophyll-a, blue-green algae, turbidity, and temperature (citation). Machine learning techniques have proven to be a powerful tool in this domain, facilitating the development of data-driven models which map observed reflectance to parameters of interest. However, these methods rely on the availability of significant volumes of in-situ reference data needed to calibrate models and asses their uncertainty. For example Aurin et al... Similarly, Ross et al developed... The challenge posed by in-situ data collection together with long satellite revisit times limits the utility of remote sensing imagery for real-time applications.

@@

Unsupervised classification

endmember extraction, spectral unmixing

@@

>>>>

Inland water bodies pose a unique challenge for analysis with remote sensing imagery due to their complex spectral characteristics and small-scale spatial variability; nevertheless, significant effort has driven the development of techniques for estimating optically-active water quality parameters such as chlorophyll-a, blue-green algae, turbidity, and temperature \cite{ritchie2003remote, bonansea2015using}. Machine learning has proven to be a powerful tool in this domain, with tree-based models, support vector machines, deep neural networks, and others demonstrating success \cite{thenkabail2018hyperspectral, sagan2020monitoring}

facilitating the development of data-driven models mapping observed reflectance to parameters of interest.

However, these methods rely on the availability of significant volumes of in-situ reference data needed to calibrate models and asses their uncertainty. For example Aurin et al... Similarly, Ross et al developed... The challenge posed by in-situ data collection together with long satellite revisit times limits the utility of remote sensing imagery for real-time applications.

- Remote sensing used for oil spill analysis: extent and thickness mapping \cite{kokaly2013spectroscopic, leifer2012state}

- Sun glitter remains a key challenge for sensing in the visible portion of the spectrum but multi and hyperspectral imagers have been used for oil spill identification and to identify their impacts on vegetation stress and mortality \cite{fingas2014review, khan2018modern}

- Combining spectral indices such as the NDVI together with machine learning is a popular approach \cite{thenkabail2018hyperspectral, sagan2020monitoring, lu2021retrieval}

- Ross et al. created a comprehensive dataset with over 600,000 water quality records matching optically active water quality parameters with associated satellite imagery from Landsat 5,7,8. To achieve this quantity of data, they needed records spanning 1984 to 2019 \cite{ross2019aquasat}
- Aurin et al. took a similar route combining 30 years of remote sensing imagery with in-situ data from over 500 field compaigns for CDOM, organic carbon, etc... \cite{aurin2018remote}.

- When ground truth data aren't available in sufficient quantity, classification into water quality categories is another approach (ground truth can be easier to obtain by expert analysis of scene) \cite{koponen2002lake}

@@

Unsupervised classification
- When no ground-truth data are available, unsupervised classification can still help partition imagery into groups or clusters.
- Many data-driven, ML methods have been employed for the task including various matrix factorizations, k-nearest neighbors, fuzzy c-means, density estimation methods, etc. \cite{zhang2019hyperspectral}
- Additionally unsupervised approaches can be used to perform nonlinear dimensionality reduction & pre-processing for supervised approaches.
- Many unsupervised approaches can be challenging to interpret due to the lack of a relationship between the identified classes, e.g. is class 1 more similar to class 2 than class 10?
- The Self Organizing Map is a popular technique which classifies data while enforcing topological relationships between the resulting classes such that classes close to each  other in the latent space are more similar than classes that are far apart.
- SOM used for remote sensing imagery classification \cite{msom-remote-sensing}
- SOM used for clustering and data compression of HSI cube-sat \cite{danielsen2021self}
- SOM used for land-use and land-cover change analysis \cite{penfound2021analysis}
- SOM has bee nused for identifying synoptic-scale patterns in wind and sea surface temperature data {som-satellite}

endmember extraction, spectral unmixing
- This is a related problem where the goal is to identify unique spectral signatures which combine (linearly or non-linearly) to produce the measured signal.
- Having identified the set of endmembers, we then seek to determine their relative abundance in each pixel
- Many statistical approaches for extracting endmembers \cite{berman2004ice}
- Sparse PCA used for endmember extraction \cite{yousefi2016mineral}
- Popular spectral unmixing approach is to treat "pure" endmembers as vertices of a simplex \cite{plaza2012endmember, nascimento2005vertex}
- Convolutional Neural Networks with Autoencoder architectures are a popular ML approach which identify endmembers and perform non-linear unmixing \cite{palsson2020convolutional, non-negative-autoencoders, su2019daen, borsoi2019deep}
- Self Organizing Map is another approach which has been used for endmember extraction together with neural networks for abundance mapping (unmixing) \cite{cantero2004analysis}

@@

- applications include food quality & safety, medical diagnoses, precision agriculture, and forensic document examination \cite{khan2018modern}
- Hyperspectral data were used to distinguish oils by type, e.g. crude, diesel, gasoline, and palm \cite{yang2020characterization}
- Near-earth HSI addresses spatial, spectral, and temporal limitations of satellite and airborne platforms. UAV-based HSI enable fine-scale mapping \cite{banerjee2020uav}
- Drones can be equipped with HSI and compute to enable rapid generation of spectral indices like the NDVI for applications such as precision agriculture \cite{horstrand2019uav}
- UAV-based HSI can be georectified to centimeter-scales without need for ground control points by using on-board GPS and IMU \cite{arroyo2019implementation}
- UAV-based HSI for turbidity estimation \cite{vogt2016near}
- Polynomial regression models for chlorophyll-a, turbidity \cite{zhang2022selection}
